Weighted Model Counting with Conditional Probabilities

### 2020-06-11 paper.tex
Some preliminary experimental results and reordering my notes and TODOs

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}

@@

\title{What Boolean Algebras Can Teach Us About Weighted Model Counting}
%\title{Generalising Weighted Model Counting to Boolean Algebras}
%\title{On the Limitations of Weighted Model Counting}
%\title{Weighted Model Counting/Integration from the Perspective of Boolean Algebras}
%\title{What Boolean Algebras Can Teach Us About Weighted Model Counting/Integration?}
\author{Paulius Dilkas}

@@

\section{Introduction}
% Feedback: What are the main claims, what are the main takeaways, intuitive
% [???] of theorems to follow. To do this, we appeal to algebraic constructions
% to define the main concepts for introducing measures on Boolean algebras.

% Feedback: Can you say something here about factorized vs non-factorized
% weight function definitions? That is, factorized is when w maps literals to
% R_>=0, non-factorized is when w maps models to R_>=0 and show:
% a) come up with nice example when non-factorized weights are intuitive
% c) clarify that the factorized definition have is w.r.t. models, in case some
% one gets confused [It doesn't have to be, if the BA is not free -- P.]

% Feedback 2: The paper at this stage is very technical - the danger is that
% WMC/SRL people may not be able to follow it and so would be hard to get
% accepted. Without clear target audience, [???] get work accepted. My main high
% level suggestion is that let us tease apart what you have and see if a story
% emerges. That is, let us attempt to write a paper *with examples* and see
% whether with significant motivation, we have a story emerging. Below, sample
% text needed to adequately motivate [???] for WMC/SRL community.

% Feedback from the panel: Clear up questions about independence between
% literals.

% TODO for later:
% 1) make up my mind about a,b vs x,y and stick to it (maybe x,y?)
% 2) Coproducts/pushouts vs (amalgamated) free products. Probably the former.
% 3) 'with generating set S' -> 'over S'

@@

@@

% TODO
% 1) Feedback 2: Explain with examples: models = elements [atoms] of algebra
% 2) Perhaps reorder the section of preliminaries into paragraphs, i.e., a
% paragraph for order, for homomorphisms, etc. This would take up less space.
% 3) how to compute the number of elements in the algebra.

@@

\section{WMC as a Measure}
% TODO: Feedback 2: Need to explain how WMC and NWMC connects to standard
% definitions of WMC and Pr(phi|e) = WMC(phi /\ e)/WMC(e).

@@

  % TODO: be careful about mentioning ideals, filters, and quotients
  The corresponding BA $B(\Delta)$ can then be constructed using
  \cref{def:algebra_from_logic}. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$ and then take a quotient with respect to
  either the filter generated by $p$ or the ideal\footnote{More details on these
    concepts can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$.

  Each element of $B(\mathcal{L})$ can also be seen as a subset of the set of
  all models of $\mathcal{L}$, with $0$ representing $\emptyset$, $1$
  representing the set of all (four) models, each atom representing a single
  model, and each edge going upward representing a subset relation. Thus,
  the Boolean-algebraic way of calculating the WMC of $\Delta$ consists of:
  \begin{enumerate} % TODO: why is step 1 always possible?
  \item Identifying an element $a \in B(\mathcal{L})$ that corresponds to
    $\Delta$.
  \item Finding all atoms of $B(\mathcal{L})$ that are `dominated' by $a$
    according to the partial order.
  \item Using $w$ to calculate the weight of each such atom.
  \item Adding the weights of these atoms.
  \end{enumerate}
  This motivates the following definition of WMC generalised to BAs.
\end{example}
% TODO: clarify what B(L) means. And whether B(Delta) is even necessary.
% TODO: reference for the set/subset thing.

% TODO: Feedback 2: You need to explain what precisely these mean in logic and
% models and weight functions are usually defined and understood.
% TODO: This should be replaced with inner sums (a.k.a. free products)
\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be an atomic BA, and let $M \subset \mathbf{B}$ be its set of
  atoms. Let $L \subset \mathbf{B}$ be such that every atom $m \in M$ can be
  uniquely expressed as $m = \bigwedge L'$ for some $L' \subseteq L$, and let
  $w\colon L \to \mathbb{R}_{\ge 0}$ be arbitrary. The \emph{weighted model
    count} $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC_w(x) = \begin{cases}
      0 & \text{if } x = 0 \\
      \prod_{l \in L'} w(l) & \text{if } M \ni x = \bigwedge L' \\
      \sum_{\text{atoms } a \le x} \WMC_w(a) & \text{otherwise}
    \end{cases}
  \]
  for any $x \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC_w\colon \mathbf{B} \to [0, 1]$ as $\nWMC_w(x) =
  \frac{\WMC_w(x)}{\WMC_w(1)}$ for all $x \in \mathbf{B}$. For both $\WMC_w$ and
  $\nWMC_w$, we will drop the subscript when doing so results in no potential
  confusion. Finally, we say that a measure $m\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$ is a \emph{WMC measure} (or is \emph{WMC-definable}) if
  there exists a subset $L \subset \mathbf{B}$ and a weight function $w\colon L
  \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$.
\end{definition}
% TODO: mention that the definition can be reduced to a single formula (i.e.,
% without cases)
% TODO: any measure is a WMC measure if all atoms are in L

@@

% TODO: Proofs need to be updated and propositions could be phrased in a better
% way, but the gist should be the same.

@@

% TODO: a special case for weight=0.
\begin{theorem}
  Let $\mathbf{B}$ be a free BA over $\{ l_i \}_{i=1}^n$ (for some $n \in
  \mathbb{N}$) with measure $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$, and let
  $L = \{ l_i \}_{i = 1}^n \cup \{ \neg l_i \}_{i = 1}^n$. Then there exists a
  weight function $w\colon L \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$ if
  and only if
  \begin{equation} \label{eq:wmccondition}
  m(l \land l') = m(l)m(l')
  \end{equation}
  for all distinct $l, l' \in L$ such that $l \ne \neg l'$.
\end{theorem}

@@

% TODO: Feedback 2: If you prove (b) above, you could motivate why weights on
% literals is attractive and whether there is a way to augment the expressivity
% of WMC while still maintaining literal level weights. Hence this action.

@@

% TODO: If L denotes literals, then it doesn't denote a generating set
\begin{theorem} \label{thm:extension}
  Let $\mathbf{B}$ be a free BA over a finite set $S$, and let $m\colon
  \mathbf{B} \to \mathbb{R}_{\ge 0}$ be an arbitrary measure. Let $L = \{ s \mid
  s \in S \} \cup \{ \neg s \mid s \in S \}$. By \cref{lemma:finite_and_free},
  we know that $\mathbf{B}$ has $n = 2^{|S|}$ atoms. Let $\{a_i\}_{i=1}^n$
  denote those atoms in some arbitrary order. Let $L' = L \cup \{ \phi_i
  \}_{i=1}^n \cup \{\neg \phi_i \}_{i=1}^n$ be the set $L$ extended with $2n$
  new elements. Let $\mathbf{B'}$ be the unique Boolean algebra with $\{ \phi_i
  \land a_i \}_{i=1}^n \cup \{ \neg \phi_i \land a_i \}_{i=1}^n$ as its set of
  atoms. Let $\iota\colon \mathbf{B} \hookrightarrow \mathbf{B'}$ be the
  inclusion homomorphism. Let $w\colon L' \to \mathbb{R}_{\ge 0}$ be defined by
  \[
    w(l) = \begin{cases}
      m(a_i)/2 & \text{if } l = \phi_i \text{ or } l = \neg\phi_i \text{ for
        some } i \in [n] \\
      1 & \text{otherwise}
    \end{cases}
  \]
  for all $l \in L'$, and note that this defines a WMC measure $m' =
  \WMC_{w}\colon \mathbf{B'} \to \mathbb{R}_{\ge 0}$. Then
  \[
    m(a) = (m' \circ \iota)(a)
  \]
  for all $a \in \mathbf{B}$.
\end{theorem}

@@

% TODO: make J depend on i
\begin{proof} % TODO: find a reference for this first claim
  Since $\mathbf{B}$ is free over $S$, each atom $a_i \in \mathbf{B}$ is an
  infimum of elements in $L$, i.e.,
  \[
    a_i = \bigwedge_{j \in J} a_{i,j}
  \]
  for some $\{ a_{i,j} \}_{j \in J} \subset L$. Moreover, each atom $b \in
  \mathbf{B'}$ can be represented as either $b = \phi_i \land a_i$ or $b =
  \neg\phi_i \land a_i$ for some atom $a_i \in \mathbf{B}$, also making it an
  infimum over a subset of $L'$. Then, for any $b \in \mathbf{B}$,
  \[
    (m' \circ \iota)(b) = \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\
        \phi_i \land a_i \le \iota(b)}} (w(\phi_i) + w(\neg\phi_i)) \prod_{j \in
    J} w(a_{i,j}),
  \]
  recognising that, for any $\iota(b)$, any atom $a_i \in \mathbf{B}$ satisfies
  $\phi_i \land a_i \le \iota(b)$ if and only if it satisfies $\neg\phi_i \land
  a_i \le \iota(b)$. Then, according to the definition of $w$,
  \[
    (m' \circ \iota)(b) = \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\
        \phi_i \land a_i \le \iota(b)}} (w(\phi_i) + w(\neg\phi_i)) =
    \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\ \phi_i \land a_i \le
        \iota(b)}} m(a_i) = m(b),
  \]
  provided that
  \[
    \phi_i \land a_i \le \iota(b) \quad \text{if and only if} \quad a_i \le b,
  \]
  but this is equivalent to
  \[
    \phi_i \land a_i = \phi_i \land a_i \land b \quad \text{if and only if}
    \quad a_i = a_i \land b
  \]
  which is true because $\phi_i \not\in L$.
\end{proof}

@@

% TODO: Feedback 2: You need to explain the significance of this result.

@@

% TODO: make sure that this is enough to guarantee a pushout
\begin{theorem}
  Let $\{ S_i \}_{i=0}^n$ be a finite set of finite sets for some $n > 1$ such
  that for all distinct positive integers $i$ and $j$, $S_i \cap S_j = S_0$, and
  let
  \[
    \mathbf{A} = \bigoplus_{\substack{\mathcal{F}(S_0)\\ 1 \le i \le n}}
    \mathcal{F}(S_i).
  \]
  Let $(m_i\colon \mathcal{F}(S_i) \to \mathbb{R}_{\ge 0})_{i=1}^n$ be arbitrary
  measures. Then there is a unique measure $m\colon \mathbf{A} \to
  \mathbb{R}_{\ge 0}$ such that, for any element $b \in \mathcal{F}(S_0)$,
  subset $J \subseteq \{ 1, 2, \dots, n \}$, and elements $\{ a_j \in
  \mathcal{F}(S_j \setminus S_0) \}_{j \in J}$,
  \[
    m \left(b \land \bigwedge_{j \in J} a_j \right) = \prod_{j \in J} m_j(b
    \land a_j).
  \]
\end{theorem}
\begin{proof}
  % TODO: start here
\end{proof}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{soul} % temp

@@

\title{Weighted Model Counting with Conditional Probabilities}
\author{Paulius Dilkas}

@@

\section{Introduction}

\paragraph{Legend.}
\begin{description}
\item[F] Feedback (round 1).
\item[F2] Feedback (round 2).
\item[FP] Feedback from the review panel.
\item[\textbullet] My own idea.
\end{description}

\paragraph{The big TODO list.}
\begin{itemize}
\item Look into ways to theoretically justify the performance benefits. Some
  suggestions:
  \begin{itemize}
  \item the number of variables,
  \item the number of clauses/CPTs/lines in CPTs,
  \item the proportion of edges that exist,
  \item average degree,
  \item average number of clauses that a variable is in,
  \item the number of clusters in the linear/tree-like decomposition.
  \end{itemize}
\item Have an example of how the ADDs function in this situation. If not for the
  paper, then at least for slides. Use the framework to check its correctness.
\item Along with comparing the (3?) encodings, I could also compare this
  to using ADDs directly on Bayesian networks.
\item[F] Give a concrete example of something impossible to represent using WMC.
\item Describe the changes to the Cachet DIMACS CNF format (and mention that, if
  needed, the assumption that $\Pr(A) + \Pr(\neg A) = 1$ can be discarded).
  Instead of \texttt{w 1 0.1}, we have \texttt{w 1 2 0.1} and \texttt{w 1 -2
    0.5}. This provides half of the information; the remaining half can be
  deduced, but the algorithm itself can support weights that are not necessarily
  probabilities.
\item Rename the repository into `conditional-wmc'.
\item Give some examples (and/or a proof) of why this is correct when combined
  with arbitrary clauses (i.e., conjunctions of disjunctions).
\item Turn the saved PDFs into citations.
\item Add a mini theorem: my model can express any probability distribution on
  the BA (just like BNs can express any discrete distribution).
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow. To do this, we appeal to algebraic constructions to
  define the main concepts for introducing measures on Boolean algebras.
\item[F] Can you say something here about factorized vs non-factorized
  weight function definitions? That is, factorized is when $w$ maps literals to
  $R_{\ge 0}$, non-factorized is when $w$ maps models to $R_{\ge 0}$ and
  \begin{itemize}
  \item come up with nice example when non-factorized weights are intuitive;
  \item clarify that the factorized definition have is w.r.t. models, in case some
    one gets confused. [It doesn't have to be, if the BA is not free---P.]
  \end{itemize}
\item[F2] The paper at this stage is very technical---the danger is that WMC/SRL
  people may not be able to follow it and so would be hard to get accepted.
  Without clear target audience, [???] get work accepted. My main high level
  suggestion is that let us tease apart what you have and see if a story
  emerges. That is, let us attempt to write a paper {\bf with examples} and
  see whether with significant motivation, we have a story emerging. Below,
  sample text needed to adequately motivate [???] for WMC/SRL community.
\item[FP] Clear up questions about independence between literals.
\item[F2] Preliminaries: explain with examples: models $=$ elements [atoms] of
  algebra.
\item WMC as a Measure
  \begin{itemize}
  \item[F2] Need to explain how WMC and NWMC connects to standard definitions of
    WMC and $\Pr(\phi \mid e) = \WMC(\phi \land e)/\WMC(e)$.
  \item[F2] You need to explain what precisely these mean in logic and models
    and weight functions are usually defined and understood.
  \item Be careful about mentioning ideals, filters, and quotients.
  \end{itemize}
\item What Measures are WMC-Definable: proofs need to be updated and
  propositions could be phrased in a better way, but the gist should be the
  same.
\item Extending the Algebra.
  \begin{itemize}
  \item[F2] If you prove (b) above, you could motivate why weights on literals
    is attractive and whether there is a way to augment the expressivity of WMC
    while still maintaining literal level weights. Hence this action.
  \item[F2] You need to explain the significance of this result.
  \end{itemize}
\item Make up my mind about $a, b$ vs. $x, y$ and stick to it (maybe $x, y$?).
\item Decide on terminology: coproducts/pushouts vs (amalgamated) free products.
  Probably the former.
\item Terminology: `with generating set $S$' $\to$ `over $S$'.
\item Perhaps reorder the section of preliminaries into paragraphs, i.e., a
  paragraph for order, for homomorphisms, etc. This would take up less space.
\item Notation: if $L$ denotes literals, then it doesn't denote a generating
  set.
\end{itemize}

\paragraph{Experimental setup.}
\begin{itemize}
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  (highlighted the initial setup, can expand later):
  \begin{itemize}
  \item \hl{ADDMC} \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Encodings:
  \begin{itemize}
  \item Darwiche \cite{DBLP:conf/kr/Darwiche02}
  \item \hl{Sang et al.} \cite{DBLP:conf/aaai/SangBK05}
  \item \hl{mine}
  \end{itemize}
\item Datasets\footnote{There might be more at
    \url{https://meelgroup.github.io/}.}
  \begin{itemize}
  \item binary Bayesian networks from Sang et
    al.\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}
    \cite{DBLP:conf/aaai/SangBK05}
    \begin{itemize}
    \item \hl{Grid networks}: computing marginal probability of the last node in
      the grid.
    \item Plan recognition problems.
    \item Deterministic quick medical reference.
    \end{itemize}
    For the last two, the task can be to calculate the marginal probability of a
    randomly chosen node (so, repeat at least ten times).
  \item non-binary Bayesian networks?
  \item ProbLog \cite{DBLP:conf/uai/FierensBTGR11}
  \item probabilistic programs \cite{DBLP:journals/corr/abs-2005-09089}
  \end{itemize}
\end{itemize}

@@

\paragraph{Notes.}
\begin{itemize}
\item Thesis: many important computational problems are solved by encoding them
  as WMC. But if the weight function is extended to `conditional probabilities',
  the problem becomes easier because current approaches need workarounds in
  order to be independent.
\item Shorter thesis: many important problems are encoded as WMC, but they can
  be encoded in a better way if we allow for conditional probabilities.
\item We extend the Gaifman graph to add edges when two variables occur in the
  same CPT (e.g., including the edge from $A$ to $B$ when the CPT is $\Pr(A \mid
  B)$).
\item Extra benefit: one does not need to come up with a way to turn some probability
distribution to into a fully independent one.
\item Trivial CPTs such as $\Pr(A \mid B) = \Pr(\neg A \mid B) = 0.5$, the ADDs
  of which simplify to a single number, are put in the first cluster.
\item Observation: by inspecting the BN, we could identify CPTs that could be
  completely ignored, but maybe good heuristics take care of that anyway.
\item Observation: ADDs have a restrict command. I'm not going to use it, but it
  could be nice.
\item Alternative scenario to consider for future work: \#\SAT{} solver
  generates solutions which are then used to restrict the ADDs that calculate
  probabilities.
\item Important future work: replacing ADDs with
  AADDs\footnote{\url{https://github.com/ssanner/dd-inference}} is likely to
  bring performance benefits. Other extensions:
  \begin{itemize}
  \item FOADDs can represent first order statements;
  \item XADDs can replace WMI for continuous variables;
  \item ADDs with intervals can do approximations.
  \end{itemize}
\end{itemize}

@@

@@

\section{WMC as a Measure}

@@

  The corresponding BA $B(\Delta)$ can then be constructed using
  \cref{def:algebra_from_logic}. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$ and then take a quotient with respect to
  either the filter generated by $p$ or the ideal\footnote{More details on these
    concepts can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$.

  Each element of $B(\mathcal{L})$ can also be seen as a subset of the set of
  all models of $\mathcal{L}$, with $0$ representing $\emptyset$, $1$
  representing the set of all (four) models, each atom representing a single
  model, and each edge going upward representing a subset relation. Thus,
  the Boolean-algebraic way of calculating the WMC of $\Delta$ consists of:
  \begin{enumerate}
  \item Identifying an element $a \in B(\mathcal{L})$ that corresponds to
    $\Delta$.
  \item Finding all atoms of $B(\mathcal{L})$ that are `dominated' by $a$
    according to the partial order.
  \item Using $w$ to calculate the weight of each such atom.
  \item Adding the weights of these atoms.
  \end{enumerate}
  This motivates the following definition of WMC generalised to BAs.
\end{example}
\todo[inline,caption={}]{
  \begin{itemize}
  \item Why is Step 1 always possible?
  \item Clarify what $B(L)$ means and whether $B(\Delta)$ is even necessary.
  \item Find a reference for the set/subset thing.
  \item This should be replaced with inner sums (a.k.a. free products).
  \item Mention that the subsequent definition can be reduced to a single
    formula (i.e., without cases).
  \item Any measure is a WMC measure if all atoms are in $L$.
  \end{itemize}
}

\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be an atomic BA, and let $M \subset \mathbf{B}$ be its set of
  atoms. Let $L \subset \mathbf{B}$ be such that every atom $m \in M$ can be
  uniquely expressed as $m = \bigwedge L'$ for some $L' \subseteq L$, and let
  $w\colon L \to \mathbb{R}_{\ge 0}$ be arbitrary. The \emph{weighted model
    count} $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC_w(x) = \begin{cases}
      0 & \text{if } x = 0 \\
      \prod_{l \in L'} w(l) & \text{if } M \ni x = \bigwedge L' \\
      \sum_{\text{atoms } a \le x} \WMC_w(a) & \text{otherwise}
    \end{cases}
  \]
  for any $x \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC_w\colon \mathbf{B} \to [0, 1]$ as $\nWMC_w(x) =
  \frac{\WMC_w(x)}{\WMC_w(1)}$ for all $x \in \mathbf{B}$. For both $\WMC_w$ and
  $\nWMC_w$, we will drop the subscript when doing so results in no potential
  confusion. Finally, we say that a measure $m\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$ is a \emph{WMC measure} (or is \emph{WMC-definable}) if
  there exists a subset $L \subset \mathbf{B}$ and a weight function $w\colon L
  \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$.
\end{definition}

@@

@@

\todo[inline]{This theorem needs a special case for zero weights.}
\begin{theorem}
  Let $\mathbf{B}$ be a free BA over $\{ l_i \}_{i=1}^n$ (for some $n \in
  \mathbb{N}$) with measure $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$, and let
  $L = \{ l_i \}_{i = 1}^n \cup \{ \neg l_i \}_{i = 1}^n$. Then there exists a
  weight function $w\colon L \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$ if
  and only if
  \begin{equation} \label{eq:wmccondition}
  m(l \land l') = m(l)m(l')
  \end{equation}
  for all distinct $l, l' \in L$ such that $l \ne \neg l'$.
\end{theorem}

@@

@@

\begin{theorem} \label{thm:extension}
  Let $\mathbf{B}$ be a free BA over a finite set $S$, and let $m\colon
  \mathbf{B} \to \mathbb{R}_{\ge 0}$ be an arbitrary measure. Let $L = \{ s \mid
  s \in S \} \cup \{ \neg s \mid s \in S \}$. By \cref{lemma:finite_and_free},
  we know that $\mathbf{B}$ has $n = 2^{|S|}$ atoms. Let $\{a_i\}_{i=1}^n$
  denote those atoms in some arbitrary order. Let $L' = L \cup \{ \phi_i
  \}_{i=1}^n \cup \{\neg \phi_i \}_{i=1}^n$ be the set $L$ extended with $2n$
  new elements. Let $\mathbf{B'}$ be the unique Boolean algebra with $\{ \phi_i
  \land a_i \}_{i=1}^n \cup \{ \neg \phi_i \land a_i \}_{i=1}^n$ as its set of
  atoms. Let $\iota\colon \mathbf{B} \hookrightarrow \mathbf{B'}$ be the
  inclusion homomorphism. Let $w\colon L' \to \mathbb{R}_{\ge 0}$ be defined by
  \[
    w(l) = \begin{cases}
      m(a_i)/2 & \text{if } l = \phi_i \text{ or } l = \neg\phi_i \text{ for
        some } i \in [n] \\
      1 & \text{otherwise}
    \end{cases}
  \]
  for all $l \in L'$, and note that this defines a WMC measure $m' =
  \WMC_{w}\colon \mathbf{B'} \to \mathbb{R}_{\ge 0}$. Then
  \[
    m(a) = (m' \circ \iota)(a)
  \]
  for all $a \in \mathbf{B}$.
\end{theorem}

@@

\todo[inline,caption={}]{
  \begin{itemize}
  \item Make $J$ depend on $i$.
  \item Find a reference for this first claim in the following proof.
  \end{itemize}
}
\begin{proof}
  Since $\mathbf{B}$ is free over $S$, each atom $a_i \in \mathbf{B}$ is an
  infimum of elements in $L$, i.e.,
  \[
    a_i = \bigwedge_{j \in J} a_{i,j}
  \]
  for some $\{ a_{i,j} \}_{j \in J} \subset L$. Moreover, each atom $b \in
  \mathbf{B'}$ can be represented as either $b = \phi_i \land a_i$ or $b =
  \neg\phi_i \land a_i$ for some atom $a_i \in \mathbf{B}$, also making it an
  infimum over a subset of $L'$. Then, for any $b \in \mathbf{B}$,
  \[
    (m' \circ \iota)(b) = \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\
        \phi_i \land a_i \le \iota(b)}} (w(\phi_i) + w(\neg\phi_i)) \prod_{j \in
    J} w(a_{i,j}),
  \]
  recognising that, for any $\iota(b)$, any atom $a_i \in \mathbf{B}$ satisfies
  $\phi_i \land a_i \le \iota(b)$ if and only if it satisfies $\neg\phi_i \land
  a_i \le \iota(b)$. Then, according to the definition of $w$,
  \[
    (m' \circ \iota)(b) = \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\
        \phi_i \land a_i \le \iota(b)}} (w(\phi_i) + w(\neg\phi_i)) =
    \sum_{\substack{\text{atoms } a_i \in \mathbf{B}:\\ \phi_i \land a_i \le
        \iota(b)}} m(a_i) = m(b),
  \]
  provided that
  \[
    \phi_i \land a_i \le \iota(b) \quad \text{if and only if} \quad a_i \le b,
  \]
  but this is equivalent to
  \[
    \phi_i \land a_i = \phi_i \land a_i \land b \quad \text{if and only if}
    \quad a_i = a_i \land b
  \]
  which is true because $\phi_i \not\in L$.
\end{proof}

@@

@@

\todo[inline]{Make sure that this is enough to guarantee a pushout.}
\begin{theorem}
  Let $\{ S_i \}_{i=0}^n$ be a finite set of finite sets for some $n > 1$ such
  that for all distinct positive integers $i$ and $j$, $S_i \cap S_j = S_0$, and
  let
  \[
    \mathbf{A} = \bigoplus_{\substack{\mathcal{F}(S_0)\\ 1 \le i \le n}}
    \mathcal{F}(S_i).
  \]
  Let $(m_i\colon \mathcal{F}(S_i) \to \mathbb{R}_{\ge 0})_{i=1}^n$ be arbitrary
  measures. Then there is a unique measure $m\colon \mathbf{A} \to
  \mathbb{R}_{\ge 0}$ such that, for any element $b \in \mathcal{F}(S_0)$,
  subset $J \subseteq \{ 1, 2, \dots, n \}$, and elements $\{ a_j \in
  \mathcal{F}(S_j \setminus S_0) \}_{j \in J}$,
  \[
    m \left(b \land \bigwedge_{j \in J} a_j \right) = \prod_{j \in J} m_j(b
    \land a_j).
  \]
\end{theorem}
